Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-08 edubots_surveys.Rnw
conclusions - par 1-2

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education can facilitate the delivery of personalized learning
that occurs anytime anywhere, promote collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries, the most popular MIM platform
used by educators to \textcolor{black}{provide} assignments' feedback to students, support course
discussions, and provide learning resources in informal learning settings
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}; however, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM in higher education is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the
use of chatbots, which, being conversational agents, usually dwell in systems
where synchronous conversations take place.  The use of conversational agents
(chatbots) in higher education is still at its infancy
\cite{yang2019opportunities}. Nevertheless, recent studies examining their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat} have led to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the use of chatbots in collecting course feedback from students in higher
education improved students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using either tools
such as mobile devices or teaching strategies based on gamification
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} \textcolor{black}{adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation}.

Several higher education chatbots' evaluation studies have been undertaken. For
instance, a recent evaluation review study presented by Smutny and Schreiberova
\cite{Smutny2020} examined 47 educational chatbots implemented in Facebook
Messenger with the focus to identify characteristics and quality metrics such as
language, subject matter and platform\textcolor{black}{, whereas the study undertaken by Pérez and collaborators
\cite{perez2020rediscovering} aimed to categorize educational chatbots, according
to their, purpose into service-oriented and teaching-oriented}. The first category
includes those that provide service support such as the chatbot Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}; both chatbots respond to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about timetables, grades, tutors and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer to
their frequently asked questions \textcolor{black}{and, for some time, the University of Granada had
\textcolor{black}{Elvira} as a chatbot in its main web page \cite{MOREO20129} to perform that duty.} Besides being able
to answer pre-established (frequent) questions, it did so from the website of
the University of Granada using an inset \textcolor{black}{persona, voiced by a real person, who
lip-synced the answers. The emphasis in the case of Elvira, however, was on the
authenticity of the synchronization more than in making the update of questions
answered simpler, or more interactive, or more open to the rest of the university staff, using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than the ones provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide an useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, something which is the focus of this paper}.

On the other hand, teaching-oriented chatbots are more sophisticated, as they
set personalized learning outcomes and monitor learning progress. For instance,
\cite{fernoagua2018intelligent} reported on ``eduAssistant'', a virtual teaching
assistant chatbot developed on the Telegram messaging platform. In this study,
the Telegram platform is chosen because it is easy to use, students are familiar
with its features, and it enables them to exchange messages in different formats
(text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram
could operate on all devices and operating systems. The ``eduAssistant'' chatbot
acts as an automatic agent in teacher-content-student, facilitating real-time
feedback loops and providing a personalized learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose questions to
their students and the chatbot assists those who need further help by giving
them more hints and reporting it to their educator's dashboard
\cite{fernoagua2018intelligent}. This can help educators locate those
students that need more attention and send them more educational resources
relevant to their academic attainment. \textcolor{black}{In addition, a recent study
has analysed how the use of chatbots positively affected the learning outcomes
of students in a Chinese class \cite{doi:10.1177/0735633120929622}}.

Despite its (arguably) successful implementation in different higher education
institutions, over all in the fringes of educational activity, not in the actual
student-teacher interaction, its implementation or deployment is not trivial.
Some authors, for instance \cite{sjostrom2018designing} have proposed a
conceptual architecture for the adoption of teaching-oriented chatbots in higher
education.  This conceptual architecture is based on a systematic literature
review of previous studies, examining the design of chatbots in higher
education, as well as making a content analysis of student emails and discussion
forum posts of four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasized the importance
of developing chatbots in platforms that students and educators are familiar
with and can easily access (i.e. Facebook Messenger) which was confirmed with
\cite{hobert2019you} and \cite{fernoagua2018intelligent} studies. In addition, \cite{sjostrom2018designing} argued that a
conceptualization of learners' questions could aid designers in integrating the
appropriate types of questions that the chatbots should support for different
courses.  \textcolor{black}{Some other authors}, like \cite{coronado2018cognitive}, proposed agents that
store learning materials to be provided on demand to students, whereas the
authors in \cite{crockett2017predicting} reported on tutoring systems, which can
perform initial assessments of students' understanding and provide learning
material that would advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many
studies have focused on the evaluation of technology acceptance and usability
\cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain
where, according to \cite{hobert2019you}, specific pedagogical factors
such as learning success and increased motivation are more important.  Therefore,
to develop \textcolor{black}{effective chatbots for higher education, all stakeholders'
(i.e. educators, students, institutions, etc) needs should be carefully
collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}. These needs include, but are not limited to, student's learning success due to higher motivation, but these are a posteriori effects that cannot be assessed in advance; this is why we focus in this paper on what the teachers are looking for in a TAM  (Technology Acceptance Model) and associated technologies like chatbots; both authors focus on what the teachers need in terms of the latter; this paper will focus on a wider perspective trying to ascertain what they are looking for in terms of general messaging technology. The classical literature \cite{tam91} already proposes that any adoption of technology must be tuned in to the user needs and experience. In this paper we will try to find out what those are in order to propose a successful model of adoption of chatbot technology}.

Along this line of research, our previous work \cite{MoraChatbots2021:anon}
aimed at analyzing the expectations of students on this regard; recent papers
have also analysed how the use of chatbots affected the learning outcomes of
students in a Chinese class \cite{doi:10.1177/0735633120929622}. This paper
analysed conversational chatbots in an one-to-one setting, finding that learning
really benefited from it. This is, once again, a proof of the benefit of
chatbots in certain settings; however, there are some prior experiences, as well as needs, that might
prevent the successful acceptance of the technology, and these are what we are
trying to find out in this paper, together with what kind of features would improve
its acceptance.

Besides, the present work is focused on the other key actor in this challenge,
i.e. educators. \textcolor{black}{How they accepted chatbots} has been studied very recently by Chocarro et
al. \cite{doi:10.1080/03055698.2020.1850426}, who, by analyzing surveys, created
a {\em technology acceptance model} that proposed a series of features that
would make chatbots easier to accept; among them, formality of language as well
as easiness and usefulness. The survey targeted primary and secondary education
teachers, and was also more interested in the general use of chatbots in
education, not specifically in a classroom setting, seeking educational
outcomes. However, they establish a series of results that are obviously
interesting and relevant for this work.

We will present next the methodology we have followed to find out teachers'
opinions.

@@

This study has employed a quantitative approach in order to address its research objective. A good amount of data was collected on the usage of messaging applications and chatbots by educators in universities and colleges.\footnote{In Spain, higher education institutions generally known elsewhere as colleges, are called simply high schools, and deliver
tertiary or associate degrees that do not have the consideration of university degrees.}
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.

In the {\bf first survey}, the multiple choice questions are
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that teachers would consider
useful for their teaching and the impact of COVID on the teaching practice; please check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and the other for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.}

This survey was piloted by the authors of the paper and their colleagues before using it in the
study.  All the comments collected from this trial were incorporated in the survey.
We primarily used mailing lists and Telegram groups to disseminate the survey link so that it would reach a larger number of educators. The form addressed at university teachers was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also  universities in Costa Rica and
Mexico. The tertiary (non-university/college) teachers who received the form were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of the participants (n=111) in this survey was 45-55 years old (39.4\%), while 91 teachers were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 teachers were older than 55 (12.1\%). Last, 84 teachers had a teaching experience of 16-25 years (29.8\%), 75 teachers 6-15 years of experience (26.6\%), 69 teachers had 0-5 years of experience (24.5\%) and finally, 54 teachers had more than 25 years of teaching experience (19.1\%).

@@

To avoid any kind of bias and to ensure the quality of the questionnaire, a panel board was consulted. There were three questions for each of the items of the questionnaire which were: Clarity, Importance, Adequateness.
Clarity was measured as an answer to the question: Is the item well presented and without ambiguity? \footnote{Original text: ¿El ítem está bien redactado y sin ambigüedad?} with two choices: Yes or No.
Both Importance and Adequateness were measured in a Likert scale from 1 to 4 (1 low, 4 high).

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers;
apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the mind of the
surveyed.  \textcolor{black}{In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based messaging systems integrated in another learning management system).}

This inclusion of university-provided messaging applications in
the survey might explain why there are so few, just 6.7\%, that answer they do not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
people might not use it, but use some extraofficial or opt-in channels such as the other included in the survey; at the end, between these official and extraofficial channels most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant difference. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions that can arise from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots and just a few subjects selected
the option "Others". This may suggest  that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive, and if the responders were more acquainted with the possibilities of the technology, possibly this response would have shown up many more times. Another way to look about it is
that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burn out to think about
innovation.

We can speculate that their opinions towards chatbots are informed by negative
experiences with customer support chatbots, which is what most people will have
experienced. This will probably require further study, but also will need to be
taken into account when introducing chatbot technologies in the classroom:
it will have to be as simple, and as satisfactory, as possible,
to (possibly) change this existing perception.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. When looking at the plot containing the sector differentiation
it is interesting to observe how university teachers prefer in a higher
proportion to use them for grades and FAQs in comparison with other tertiary education teachers.
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation seems more helpful.
Furthermore, the high student/teacher
ratios that university teachers have to deal with could be another reason to
have a higher number of them willing to use chatbots for FAQs. This could
ameliorate the number of misunderstandings and email overload that so many
students might generate.

The results, when grouping by experience, confirm that
experienced teachers are not afraid of changes and are willing to accept new
technological challenges; at least, in the same way that younger teachers
are. From this concrete plot it is interesting as well to observe the fact that very few
young/inexperience teachers are willing to use chatbots for other uses. This
reinforces the hypothesis thrown at the beginning of this section that younger
teachers have less experience to consider new ways of using chatbots in their daily work.

As it appears in the responses presented above, teachers’ responses are equally
distributed between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone number; that is, none of them is a
prevailing factor on choosing one technology over another. However, for the features: familiarity,
horizontally, sustainability and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the
majority of teachers seems to consider not valuable the presence of official
formation, pluggability, unidirectionality and synchrony in the interaction
media features of messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations \textcolor{black}{on data protection},
   but also to avoid the overhead required to \textcolor{black}{sign up new students every year.
   In general support from specific formation and access to IT help desk for specific features
   enhances the possibility of adoption by the teachers or the possibility of a successful
   adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by the university, and
   also they value sustainability as well as "officiality" of the platform, as shown in
   Table \ref{tab:interaction}}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

>>>>

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries, the most popular MIM platform
used by educators to provide assignments' feedback to students, support course
discussions, and contribute resources in informal learning settings
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

Several studies have been conducted to evaluate chatbots in higher education. For
instance, a recent study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, with a focus on identifying characteristics and quality metrics such as
language, subject matter, and platform. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which responds to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada had
Elvira, a built-in chatbot in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the synchronization rather than making the updated 
of answered questions simpler, more interactive, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

On the other hand, teaching-oriented chatbots are more sophisticated, as they
set personalized learning outcomes and monitor learning progress. For instance,
\cite{fernoagua2018intelligent} reported on ``eduAssistant'', a virtual teaching
assistant chatbot developed on the Telegram messaging platform. In this study,
the Telegram platform was chosen because it is easy to use, students are familiar
with its features, and it enables them to exchange messages in different formats
(text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram
can operate on all devices and operating systems. The ``eduAssistant'' chatbot
acts as an automatic agent in the teacher-content-student relationship, providing real-time
feedback loops and a personalized learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose questions to
their students and the chatbot assists those who need further help by giving
them more hints and reporting it to the educator's dashboard
\cite{fernoagua2018intelligent}. This can help educators identify
students who need more attention and provide them with more educational resources
relevant to their academic achievement. %In addition, a recent study
%has analysed how the use of chatbots positively affected the learning outcomes
%of students in a Chinese class \cite{doi:10.1177/0735633120929622}.

Despite its (arguably) successful implementation in different higher education
institutions, particularly on the fringes of educational activity rather than actual
student-teacher interaction, its implementation or deployment is not trivial.
Some authors, for instance \cite{sjostrom2018designing}, have proposed a
conceptual architecture for the adoption of teaching-oriented chatbots in higher
education.  This conceptual architecture is based on a systematic literature
review of previous studies examining the design of chatbots in higher
education, as well as making a content analysis of student emails and discussion
forum posts from four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasized the importance
of developing chatbots in platforms that students and educators are familiar
with and can easily access (i.e. Facebook Messenger), which was confirmed by
\cite{hobert2019you} and \cite{fernoagua2018intelligent}. In addition, \cite{sjostrom2018designing} argued that 
conceptualizing learners' questions could help designers integrate the
appropriate types of questions that chatbots should support for different
courses. Other authors, such as \cite{coronado2018cognitive}, have proposed agents that
store learning materials to be provided on demand to students, while \cite{crockett2017predicting} 
reported on tutoring systems that can
perform initial assessments of students' understanding and provide learning
materials to advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many
studies have focused on the evaluation of technology acceptance and usability
\cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain
where, according to \cite{hobert2019you}, specific pedagogical factors
such as learning success and increased motivation are more important.  Therefore,
to develop effective chatbots for higher education, the needs of all stakeholders
(i.e. educators, students, institutions, etc) should be carefully
collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}. These needs include, but are not limited to, student learning success due to higher motivation, but these are a posteriori effects that cannot be assessed in advance. This is why we focus in this paper on what  teachers are looking for in a Technology Acceptance Model (TAM) and associated technologies like chatbots. Both authors focus on what teachers need in terms of the latter; this paper will focus on a wider perspective, trying to determine what they are looking for in terms of general messaging technology. The classical literature \cite{tam91} already proposes that any adoption of technology must be tuned to the user needs and experience. In this paper, we will try to find out what those are in order to propose a successful model of adoption of chatbot technology.

In line with this research, our previous work \cite{MoraChatbots2021:anon}
aimed at analyzing the expectations of students in this regard. Recent papers
have also analysed how the use of chatbots affected the learning outcomes of
students in a Chinese class \cite{doi:10.1177/0735633120929622}. This paper
analysed conversational chatbots in a one-to-one setting, finding that learning
truly benefited from it. This is, once again, an evidence of the benefit of
chatbots in certain settings; however, there are some prior experiences, as well as needs, that may
prevent the successful acceptance of the technology, and these are what we are
trying to find out in this paper, alng with what kind of features would improve
its acceptance.

Besides, this work is focused on the other key actor in this challenge: 
educators. How they accept chatbots has been studied very recently by Chocarro et
al. \cite{doi:10.1080/03055698.2020.1850426}, who, by analyzing surveys, created
a TAM that proposed a series of features that
would make chatbots easier to accept, including formality of language as well
as easiness and usefulness. The survey targeted primary and secondary education
teachers and was also more interested in the general use of chatbots in
education, not specifically in a classroom setting, seeking educational
outcomes. However, their results are obviously
interesting and relevant for this work.

We will present next the methodology we followed to find out teachers' opinions.

@@

\textcolor{black}{This study followed a quantitative approach towards addressing its research objective, by collecting rich data about the use of messaging applications and chatbots
by educators in both universities and colleges\footnote{In Spain, they are called simply high schools, and deliver
tertiary or associated degrees that do not have the consideration of university degrees}.
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.}

In the {\bf first survey}, the multiple choice questions are
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that teachers consider
useful for their teaching and the impact of COVID on the teaching practice; please check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and the other for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.}

\textcolor{black}{This survey was piloted by the authors of the paper and their colleagues before using it in the
study.  All the comments collected from this trial were incorporated in the survey.
We mainly used mailing lists, as well as Telegram groups,
to post the survey link to reach out to more educators. The university form was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also  universities in Costa Rica and
Mexico. The tertiary (non-university/college) teachers were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.}

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). \textcolor{black}{From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of the participants (n=111) in this survey was 45-55 years old (39.4\%), while 91 teachers were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 teachers were older than 55 (12.1\%). Last, 84 teachers had a teaching experience of 16-25 years (29.8\%), 75 teachers 6-15 years of experience (26.6\%), 69 teachers had 0-5 years of experience (24.5\%) and finally, 54 teachers had more than 25 years of teaching experience (19.1\%).}

@@

To avoid any bias and to be sure about the quality of the questionnaire, a panel board was consulted. There were three questions for each of the items of the questionnaire which were: Clarity, Importance, Adequateness.
Clarity was measured as an answer to the question: Is the item well presented and without ambiguity? \footnote{Original text: ¿El ítem está bien redactado y sin ambigüedad?} with two choices: Yes or No.
Both Importance and Adequateness were measured in a Likert scale from 1 to 4 (1 low, 4 high).

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

This inclusion of university-provided messaging applications in
the survey might explain why there are so few teachers, just 6.7\%, that answered they did not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
teachers might not use it, but use some extraofficial or opt-in channels such as the others included in the survey; in the end, between these official and extra-official channels, most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions arising from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots, and just a few subjects selected
the option "Others". This may suggest that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive. If the responders were more acquainted with the possibilities of this technology, this response would possibly have shown up many more times. Another way to look at it is that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burnt out to think about innovation.

We can speculate that teachers' opinions of chatbots are influenced by negative
experiences with customer support chatbots, which is what most people are likely to have
experienced. Further study may be needed to confirm this, but it is important to 
keep in mind when introducing chatbot technologies in the classroom.
To potentially change this existing perception, it will be necessary to make the chatbot as simple and satisfying to use as possible.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. However, when we look at the plot containing the sector differentiation,
it is interesting to note that university teachers are more likely to use chatbots 
for grades and FAQs compared to other tertiary education teachers.
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation would be more helpful.
Furthermore, the high student-to-teacher
ratios that university teachers often face could be another reason why
a higher number of them are willing to use chatbots for FAQs, as this could
reduce misunderstandings and email overload from students.

The results, when grouping by teaching experience, confirm that
experienced teachers are just as open to change and willing to embrace new
technological challenges as younger teachers. 
From this concrete plot it is interesting as well to observe the fact that very few
young teachers are willing to use chatbots for other uses. This
reinforces the hypothesis thrown at the beginning of this section that younger
teachers have less experience to consider new ways of using chatbots in their daily work.

According to the responses presented above, teachers’ responses are evenly
split between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone number. This means that none of these features is a
deciding factor in their choice of technology. However, for features such as familiarity,
horizontality, sustainability, and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the majority of teachers
do not seem to see the media features of official certification, pluggability, unidirectionality, 
and synchrony as valuable in messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their institutions. This
   is probably due to the strict European data regulations on data protection,
   but also to avoid the overhead required to sign up new students every year.
   In general, additional support with specific formation and access to IT help desk for specific features
   enhance the possibility of a successful adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by their university, and
   they the "officiality" of the platform, as shown in
   Table \ref{tab:interaction}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years. Therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, teachers want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only 2.4\% of teachers are requesting more features. This could be a sign of technological burnt out produced by the previous pandemic courses, when the use of computers significantly increased.
\end{itemize}

The analysis of the questions related to RQ4 seems to indicate that teachers
prefer to share an interaction space with their students. As
shown in Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, the majority of responders
indicated that they {\em do not} want unidirectional communication, and that they prefer
{\em horizontal} interaction. This point is reinforced by results on the kind of
chatbots they preferred: most of them indicated FAQ bots, which might indicate a
need to offload part of the burden of answering every single question posed by
their students. In a horizontal setting, other students would answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
in that way, the student activity in the chat tool, answering questions, or their regularity can all be
assessed as part of the process of achieving learning objectives.
